---
title: "Coursera - Practical Machine Learning - Project"
author: "Marcos Takeshi Sato"
date: "August 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download files
We need to download the dataset files from the links below:</br>
  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv</br>
  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv</br>
  
and set the working directory ( setwd("") ) to the path of these files.

## Data
The first thing to do is load the dataset (training) and clean it.

```{r}
exercise <- read.csv("./pml-training.csv", na.strings = c("", "NA", "#DIV/0!"))
```

We can remove the columns that have too many NA values by calculating the NA ratio for each column (and remove those column with NA ratio higher than 70%)</br>
We will remove the ID column as well.</br>
*note that we search from the last column to the first to avoid problems by the removal of the current column
```{r}
for (x in length(exercise):1) {
  if (dim(exercise[is.na(exercise[,x])==TRUE,][x])[1] / nrow(exercise) >= 0.7) { 
    exercise <- exercise[,-x] 
  }
}
exercise$X <- NULL
```

Now we can set the seed for reproducible purposes:
```{r}
set.seed(777)
```

And split the training dataset for the model and the predict:
```{r}
library(caret)
inTrain <- createDataPartition(y = exercise$classe, p = 0.7, list = FALSE)
training <- exercise[inTrain,]
testing <- exercise[-inTrain,]
```

## Training
As we could see in one of the exercises of the quizzes, Random Forest method can give us a high accuracy.
```{r}
library(randomForest)
modRF <- randomForest(classe ~ ., data = training)
```
*(it took 38 seconds to process)

## Testing
With the model and the testing data, now we can predict:
```{r}
predRF <- predict(modRF, testing)
```

And test the results:
```{r}
confusionMatrix(predRF, testing$classe)
```

## Applying to other data
With the Random Forest model, now we can solve the exercise:
*we need to ajust the levels os the factor variables first.
```{r}
newExercise <- read.csv("./pml-testing.csv", na.strings = c("", "NA", "#DIV/0!"))
for (x in 1:length(newExercise)) {
  for (y in 1:length(training)) {
    if ((names(newExercise[x]) == names(training[y])) & (class(newExercise[,x]) == "factor")) {
      levels(newExercise[,x]) <- levels(training[,y])
    }
  }
}
newPred <- predict(modRF, newExercise)
```

And this is the result:
```{r}
newPred
```
